---
title: "R Notebook"
author: "Bernice Man"
date: "September 24, 2019"
output:
  word_document:
    reference_docx: word-styles-reference-01.docx
  pdf_document: default
---

Table 1 document for Analysis 1: Descriptive characteristics of sample. 

Alternative to knit: rmarkdown::render('analysis 1.rmd')

##### Table 1

```{r echo=FALSE, results="asis"}
library(arsenal)

# Read datafile, Change ChiSq test to Fisher exact, compare median WTHratio instead of mean, change to row percent
sub.data <- readRDS(file = file.choose())
#using 20210314_105719_dpp_gdm_data.Rda

# calculate expected cell numbers
chisq.test(sub.data$ASSIGN,sub.data$AGEGROUP)$expected
chisq.test(sub.data$ASSIGN,sub.data$RACE_ETH)$expected
chisq.test(sub.data$ASSIGN,sub.data$Smoke)$expected
chisq.test(sub.data$ASSIGN,sub.data$SIPCOS)$expected
chisq.test(sub.data$ASSIGN,sub.data$BMI_CAT)$expected
chisq.test(sub.data$ASSIGN,sub.data$BMIGROUP)$expected
chisq.test(sub.data$ASSIGN,sub.data$FamHx)$expected
chisq.test(sub.data$ASSIGN,sub.data$DIABF)$expected

# comparisons using fisher exact ( a few expected cells < 5) and KWT
fetable1 <- tableby(ASSIGN ~ AGEGROUP + RACE_ETH + Smoke + SIPCOS + BMI_CAT + BMIGROUP + FamHx + SIBIRTH + WaistCircAve + WTHRatio + G000 + HBA1 + totalMETs + SOSBPA + SODBPA + TRIG + DIABF, cat.test = "fe", numeric.test = "kwt", simulate.p.value = TRUE, numeric.stats = "medianq1q3" , data= sub.data)

# comparisons using chisq and KWT
table1 <- tableby(ASSIGN ~ 
                    chisq(AGEGROUP, "countpct") +
                    chisq(RACE_ETH, "countpct") +
                    chisq(Smoke, "countpct", cat.simplify = FALSE) +
                    chisq(SIPCOS, "countpct", cat.simplify = FALSE) +
                    #fisher.test(SIPCOS,) +
                    chisq(BMI_CAT, "countpct") + 
                    #fisher.test(BMI_CAT) + 
                    chisq(BMIGROUP, "countpct") +
                    #fisher.test(BMIGROUP,) +
                    chisq(FamHx, "countpct") +
                    #fisher.test(FamHx, ) + 
                    kwt(SIBIRTH, "medianq1q3", digits = 1) + 
                    kwt(WaistCircAve, "medianq1q3", digits = 1) +
                    kwt(WTHRatio, "medianq1q3", digits = 3) +
                    kwt(G000, "medianq1q3", digits = 0) + 
                    kwt(HBA1, "medianq1q3", digits = 1) + 
                    kwt(totalMETs, "medianq1q3", digits = 1) +
                    kwt(SOSBPA, "medianq1q3", digits = 0) +
                    kwt(SODBPA, "medianq1q3", digits = 0) +
                    kwt(TRIG, "medianq1q3", digits = 0) +
                    chisq(DIABF, "countpct"),
                  #fe(DIABF, "countpct", )
                    control = tableby.control(total = FALSE),
                    numeric.simplify=TRUE,
                    cat.simplify=TRUE,
                    data = sub.data)

tablabels <- list(
  AGEGROUP = "Age Group, N(%)",
  RACE_ETH = "Ethnicity, N(%)",
  Smoke = "Smoking Status, N(%)",
  SIPCOS = "PCOS History, N(%)",
  FamHx = "Family History, N(%)",
  SIBIRTH = "Live Births, Median",
  BMI_CAT = "BMI Category (kg/m\U00B2), N(%)",
  BMIGROUP = "BMI Group (kg/m\U00B2), N(%)",
  WaistCircAve = "Waist Circumference, Median",
  WTHRatio = "Waist:Hip Ratio, Median",
  G000 = "Fasting Glucose, Median",
  HBA1 = "Hemoglobin A1c, Median",
  totalMETs = "MET-hours/week, Median",
  SOSBPA = "Systolic BP (mm Hg), Median",
  SODBPA = "Diastolic BP (mm Hg), Median",
  TRIG = "Triglyceride (mg/dL), Median",
  ASSIGN = "Treatment Arm, N(%)"
)

summary(fetable1)
summary(table1, 
        title = "Baseline Characteristics by Treatment Assignment",
        labelTranslations = tablabels,
        digits.count = 0,
        digits.pct = 1,
        digits.p = 2)
```

##### Missing Data

```{r}
apply(is.na(sub.data),2,sum)
```

##### Distributions

```{r}
library(ggplot2)

ggplot(data = sub.data, aes(SIBIRTH)) +
  geom_histogram(binwidth = 1) +
  labs(x = "Live Births")

ggplot(data = sub.data, aes(WTHRatio)) +
  geom_histogram(binwidth = .01) +
  labs(x = "Waist:Hip Ratio")

ggplot(data = sub.data, aes(totalMETs)) +
  geom_histogram(binwidth = 10) +
  labs(x = "Total METs")
```

```{r echo=FALSE, results="asis"}

library(arsenal)
# table of baseline characteristics by diabetes outcome, fetable5 uses fisher exact test and median for all continuous, 

fetable5<- tableby(DIABF ~ AGEGROUP + RACE_ETH + Smoke + SIPCOS + BMI_CAT + BMIGROUP + FamHx + SIBIRTH + WaistCircAve + WTHRatio + G000 + HBA1 + totalMETs + SOSBPA + SODBPA + TRIG + ASSIGN, cat.test = "fe", numeric.test = "kwt", simulate.p.value = TRUE, numeric.stats = "medianq1q3" , data= sub.data)

summary(fetable5)


table5 <- tableby(DIABF ~ 
                    chisq(AGEGROUP, "countpct") +
                    chisq(RACE_ETH, "countpct") +
                    chisq(Smoke, "countpct", cat.simplify = FALSE) +
                    chisq(SIPCOS, "countpct", cat.simplify = FALSE) +
                    chisq(BMI_CAT, "countpct") + 
                    chisq(BMIGROUP, "countpct") +
                    chisq(FamHx, "countpct") +
                    chisq(ASSIGN, "countpct") +
                    kwt(SIBIRTH, "medianq1q3", digits = 1) + 
                    kwt(WaistCircAve, "medianq1q3", digits = 1) +
                    kwt(WTHRatio, "medianq1q3", digits = 3) +
                    kwt(G000, "medianq1q3", digits = 0) + 
                    kwt(HBA1, "medianq1q3", digits = 1) + 
                    kwt(totalMETs, "medianq1q3", digits = 1) +
                    kwt(SOSBPA, "medianq1q3", digits = 0) +
                    kwt(SODBPA, "medianq1q3", digits = 0) +
                    kwt(TRIG, "medianq1q3", digits = 0),
                    
                    control = tableby.control(total = FALSE),
                    numeric.simplify=TRUE,
                    cat.simplify=TRUE,
                    data = sub.data)



summary(table5, 
        title = "Baseline Characteristics by Diabetes Outcome",
        labelTranslations = tablabels,
        digits.count = 0,
        digits.pct = 1,
        digits.p = 2)

```

``````{r echo=FALSE, results="asis"}

library(arsenal)

# table of baseline characteristics by diabetes outcome, change mean to median for WTH ratio, 
# with row percents, fetable5 uses fisher exact test and median for all continuous, 

fetable5<- tableby(DIABF ~ AGEGROUP + RACE_ETH + Smoke + SIPCOS + BMI_CAT + BMIGROUP + FamHx + SIBIRTH + WaistCircAve + WTHRatio + G000 + HBA1 + totalMETs + SOSBPA + SODBPA + TRIG + ASSIGN, cat.test = "fe", numeric.test = "kwt", simulate.p.value = TRUE, numeric.stats = "medianq1q3" , data= sub.data)

table5 <- tableby(DIABF ~ 
                    chisq(AGEGROUP, "countrowpct") +
                    chisq(RACE_ETH, "countrowpct") +
                    chisq(Smoke, "countrowpct", cat.simplify = FALSE) +
                    chisq(SIPCOS, "countrowpct", cat.simplify = FALSE) +
                    chisq(BMI_CAT, "countrowpct") + 
                    chisq(BMIGROUP, "countrowpct") +
                    chisq(FamHx, "countrowpct") +
                    chisq(ASSIGN, "countrowpct") +
                    kwt(SIBIRTH, "medianq1q3", digits = 1) + 
                    kwt(WaistCircAve, "medianq1q3", digits = 1) +
                    kwt(WTHRatio, "medianq1q3", digits = 3) +
                    kwt(G000, "medianq1q3", digits = 0) + 
                    kwt(HBA1, "medianq1q3", digits = 1) + 
                    kwt(totalMETs, "medianq1q3", digits = 1) +
                    kwt(SOSBPA, "medianq1q3", digits = 0) +
                    kwt(SODBPA, "medianq1q3", digits = 0) +
                    kwt(TRIG, "medianq1q3", digits = 0),
                    
                    control = tableby.control(total = FALSE),
                    numeric.simplify=TRUE,
                    cat.simplify=TRUE,
                    data = sub.data)

summary(fetable5)
summary(table5, 
        title = "Baseline Characteristics by Diabetes Outcome (row percents)",
        labelTranslations = tablabels,
        digits.count = 0,
        digits.pct = 1,
        digits.p = 2)
summary (sub.data$DIABF)
```

