---
title: "R Notebook"
author: "Bernice Man"
date: "September 13, 2019"
output: 
  word_document:
    reference_docx: "word-styles-reference-01.docx"
---

Analysis 2: Cox proportional hazards model for diabetes prediction in GDM 

Alternative to knit: rmarkdown::render('analysis 2.rmd')

##### Descriptives for Variables Scaled

```{r echo=FALSE}

library(psych)
library(survival)
library(broom)
library(ggplot2)
library(Rmisc)
library(knitr)
library(tidyverse)
library(caret)
library(rms)
library(survminer)

# Read datafile
sub.data <- readRDS(file = file.choose())

# Make sure Placebo is reference group and drop 'troglitazone' (unused levels)
sub.data$ASSIGN <- relevel(sub.data$ASSIGN, ref = "Placebo")
sub.data$ASSIGN <- droplevels(sub.data$ASSIGN)

# Variables to normalize
varnorm <- c('AvgHeight', 'WTHRatio','G000','HBA1','logTRIG')
zvarnorm <- paste0('z', varnorm)

# Normalize and create new variable 'z' + old variable name, and add to data frame
sub.data[zvarnorm] <- lapply(sub.data[varnorm], scale)

describe(sub.data[, c(varnorm, zvarnorm)])

# Number of subjects before and after selecting complete cases
varselect <- c('ASSIGN','zAvgHeight','zWTHRatio','BMIGROUP','zG000','zHBA1','zlogTRIG')
cat('Number of subjects:', nrow(sub.data))
sub.data <- sub.data[complete.cases(sub.data[, varselect]), ]
cat('Number of subjects with complete variables:', nrow(sub.data), '(', varselect, ')')
```

##### Table 2 Cox Proportional Hazards Model

```{r echo=FALSE}

# Cox Proportional Hazard Model
# DIABF of No = censored (0), Yes = diabetes (1)
# New variable for DM outcome is DiabYes
sub.data <- sub.data %>%
  mutate(DiabYes = ifelse(DIABF == "Yes", 1, 0))

# Specify interactions
varint <- c('ASSIGN*BMIGROUP')

# Specify units of time (years)
units(sub.data$DIABT) <- "Year"

# Create CPH Model, DIABT is follow up time with right censored data, and DiabYes is event (1 = diabetes event)
m1surv <- Surv(sub.data$DIABT, sub.data$DiabYes, type = "right")
m1form <- as.formula(paste("m1surv ~ ", paste(c(varselect, varint), collapse = "+")))

m1 <-coxph(m1form, data=sub.data, ties='efron')
m1tidy <- tidy(m1)
m1tidy$term <- c('Lifestyle', 'Metformin', 'Height (mean), Std', 'Waist:Hip Ratio, Std',
                       'BMI 30-35', 'BMI 35+', 'Fasting Glucose, Std', 'HbA1c, Std', 'Triglycerides (ln), Std',
                       'Lifestyle * BMI 30-35', 'Metformin * BMI 30-35', 
                       'Lifestyle * BMI 35+', 'Metformin * BMI 35+')
kable(m1tidy, digits=3)
```

##### Table 3 Covariates and Proportional Hazards Assumption

```{r echo=FALSE}
# Check each covariate (proportional hazards assumption)
res1 <- cox.zph(m1)
res1

# Plot residuals (Schoenfeld residuals vs. time)
plot(res1)
```

##### Figure 1 Survival Plots

```{r echo=FALSE}

# Create new data frame and set DiabYes to 0 (no diabetes) and DIABT to 3 (3 years)
newdf <- sub.data
newdf$DIABT <- 3
newdf$DiabYes <- 0

# Calculate predicted probability of diabetes at 3 years based on cph model
sub.data$survprob <- exp(-predict(m1, newdata=newdf, type="expected"))

# Plot predicted probability by BMI graoup and treatment assignment
ggplot(sub.data, aes(y=survprob, x=BMIGROUP, fill=ASSIGN)) +
     geom_boxplot() +
     scale_y_continuous("Predicted Survival Probability at 3 years") +
     scale_x_discrete("BMI Group") +
     theme_bw()
```

##### Figure 2 Observed Vs. Predicted Probability of Diabetes

```{r echo=FALSE}

# Convert predicted probabilities to factor of deciles
sub.data$diabprobf <- factor(floor((1-sub.data$survprob)*10),
                      levels=0:9,
                      labels=c("0-9%","10-19%","20-29%","30-39%",
                              "40-49%","50-59%","60-69%","70-79%",
                              "80-89%","90-99%"))

# For each decile get std dev, std error, conf int
agg <- summarySE(sub.data, measurevar="DiabYes", groupvars="diabprobf")

# As NaNs produced (maybe only 1 case in a percentile) use only complete cases so some deciles may be missing
agg <- agg[complete.cases(agg), ]

# Create plot of observed vs. predicted probability of diabetes
ggplot(agg, aes(x=diabprobf, y=DiabYes)) +
      geom_point() +
      geom_errorbar(aes(ymin=pmax(DiabYes-ci, 0), ymax=pmin(DiabYes+ci, 1))) +
      scale_y_continuous("Observed prevalence of diabetes", breaks=seq(0, 1, .1)) +
      scale_x_discrete("Predicted 3-year probability of diabetes") +
      theme_bw()
```

##### Figure 3 Goodness of CPH Model Fit

```{r echo=FALSE}

# Check proportional hazards assumption
ggcoxdiagnostics(m1, type="schoenfeld")

# Check nonlinearity
ggcoxdiagnostics(m1, type="martingale")

# Check influential observations
ggcoxdiagnostics(m1, type="deviance")
```

##### Figure 4 Validation

```{r echo=FALSE}

# References
# https://socialsciences.mcmaster.ca/jfox/Books/Companion-1E/appendix-cox-regression.pdf

# Check for concordance (includes (1-c-statistic))
concordance(m1surv ~ predict(m1))

# Use cph from rms package (should give similar results to coxph)
m2 <- cph(formula = m1form, data = sub.data, x = TRUE, y = TRUE, surv = TRUE, time.inc = 1)
m2

# Validation with 10-fold cross validation (set seed for reproducibility)
set.seed(1)
v1 <- validate(m2, method="crossvalidation", B = 10, dxy=T)
v1

# C-statistic by Dxy
# https://www.rdocumentation.org/packages/rms/versions/5.1-3.1/topics/validate.cph
# Dxy = 2 * (C - 0.5) where C is the c-index of concordance probability
Dxy = v1[rownames(v1)=="Dxy", colnames(v1)=="index.corrected"]
orig_Dxy = v1[rownames(v1)=="Dxy", colnames(v1)=="index.orig"]
bias_corrected_c_index  <- (abs(Dxy)/2)+0.5
orig_c_index <- (abs(orig_Dxy)/2)+0.5
cat("Bias corrected c-index: ", bias_corrected_c_index)
cat("Original c-index: ", orig_c_index)

# 3-year probability estimates with calibrate
cal <- calibrate(m2, B=10, u=3)
plot(cal, ylim = c(0,1))
```

##### k-fold cross validation of logistic regression

```{r echo=FALSE}

# K-fold validation with logistic model
train_control <- trainControl(method="cv", number = 10, savePredictions = TRUE)
m3 <- train(DIABF ~ ASSIGN + zWTHRatio + BMIGROUP + zG000 + zHBA1, 
            data = sub.data, trControl = train_control, method = "glm", family = "binomial")

# Show the final model
m3$finalModel
```


