---
title: "R Notebook: Diabetes Prediction from DPP in GDM"
author: "Bernice Man"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document:
    reference_docx: word-styles-reference-01.docx
  html_document:
    df_print: paged
  pdf_document: default
---

# Analysis 2: Cox proportional hazards model for diabetes prediction in GDM 

Alternative to knit: rmarkdown::render('analysis 2.rmd')

This analysis includes the development of Cox proportional hazards model to predict diabetes in women with history of gestational diabetes. The results are submitted to publication in PLOS One.

Results include the generation of:

- Survival curves
- Tables and figures
- ROC curves

##### Descriptives for Variables Scaled

```{r load_libraries, echo=TRUE}
library(psych)
library(survival)
library(broom)
library(ggplot2)
library(Rmisc)
library(knitr)
library(tidyverse)
library(caret)
library(rms)
library(survminer)
library(e1071)
library(survivalROC)
library(plotROC)
```

```{r read_data, echo = TRUE}
# Read datafile
raw_data <- readRDS(file = file.choose())
#raw_data <- readRDS(file = "20200114_122907_dpp_gdm_data.Rda")
```

```{r drop_tzd, echo = FALSE}
# Make sure ASSIGN is a factor
# Make sure Placebo is reference group and drop 'troglitazone' (unused levels)- already dropped in main analysis
raw_data$ASSIGN <- factor(raw_data$ASSIGN)
raw_data$ASSIGN <- relevel(raw_data$ASSIGN, ref = "Placebo") 
raw_data$ASSIGN <- droplevels(raw_data$ASSIGN)
```

```{r normalize_data, echo = FALSE}
# Variables to normalize
varnorm <- c('AvgHeight', 'WTHRatio','G000','HBA1','logTRIG','WaistCircAve','totalMETs')
zvarnorm <- paste0('z', varnorm)

# Normalize and create new variable 'z' + old variable name, and add to data frame
raw_data[zvarnorm] <- lapply(raw_data[varnorm], scale)
```

```{r describe_data, echo = FALSE}
# Describe the results
psych::describe(raw_data[, c(varnorm, zvarnorm)])

# Number of subjects/rows
cat('Number of subjects:', nrow(raw_data))
```

```{r set_units, echo = FALSE}
# Specify units of time (years)
units(raw_data$DIABT) <- "Year"
```

```{r set_outcome, echo = FALSE}
# Add 0/1 outcome for diabetes event (1 = event), DIABF of No = censored (0), Yes = diabetes (1)
raw_data$DiabYes = ifelse(raw_data$DIABF == "Yes", 1, 0)
```

```{r set_seed, echo = FALSE}
# For reproducibility
set.seed(1)
```

##### Table 1 

See analysis 2.Rmd.

\newpage

```{r run_model, echo = FALSE}

# Create cph model, tidy it up, based on vector of variables, interactions, and indicate labels
run_model <- function(varselect, varint, labels) {
  # Select complete cases  
  sub_data <- raw_data[complete.cases(raw_data[, varselect]), ]
    
  cat('Number of subjects with complete variables:', nrow(sub_data), '(', varselect, ')')
  
  # Create CPH Model, DIABT is follow up time with right censored data, and DiabYes is event (1 = diabetes event)
  # Model developed from analysis of parous women with and without GDM, removed 'Height (mean), Std'
  surv <- Surv(sub_data$DIABT, sub_data$DiabYes, type = "right")
  form <- as.formula(paste("surv ~ ", paste(c(varselect, varint), collapse = "+")))
  m <- coxph(form, data = sub_data, ties = 'efron')
  
  # Exponentiate false for equation and coefficients
  mtidy <- tidy(m, exponentiate = FALSE, conf.int = TRUE, conf.level = 0.95)
  mtidy$term <- labels
  
  # Exponentiate true to see hazard ratios
  mtidy_exp <- tidy(m, exponentiate = TRUE, conf.int = TRUE, conf.level = 0.95)
  mtidy_exp$term <- labels
  
  # Return the model and package with all details in a list
  return(list(model = m, tidy_model = mtidy, tidy_model_exp = mtidy_exp, surv = surv, data = sub_data, formula = form, varselect = varselect, varint = varint, labels = labels))
}

```

```{r validate_model, echo = FALSE}

# References
# https://socialsciences.mcmaster.ca/jfox/Books/Companion-1E/appendix-cox-regression.pdf

# k-fold cross validation of logistic regression, take cph model and cross validate
validate_model <- function(m) {
  
  # Check for concordance (includes (1-c-statistic))
  surv <- m[["surv"]]
  concordance(surv ~ predict(m[["model"]]))

  # Use cph from rms package (should give similar results to coxph)
  rms_m <- cph(formula = m[["formula"]], data = m[["data"]], x = TRUE, y = TRUE, surv = TRUE, time.inc = 1)
  rms_m

  # Validation with 10-fold cross validation
  v1 <- validate(rms_m, method = "crossvalidation", B = 10, dxy = T)
  v1

  # C-statistic by Dxy
  # https://www.rdocumentation.org/packages/rms/versions/5.1-3.1/topics/validate.cph
  # Dxy = 2 * (C - 0.5) where C is the c-index of concordance probability
  Dxy = v1[rownames(v1) == "Dxy", colnames(v1) == "index.corrected"]
  orig_Dxy = v1[rownames(v1) == "Dxy", colnames(v1) == "index.orig"]
  bias_corrected_c_index <- (abs(Dxy)/2)+0.5
  orig_c_index <- (abs(orig_Dxy)/2)+0.5
  cat("Bias corrected c-index: ", bias_corrected_c_index)
  cat("Original c-index: ", orig_c_index)
  
  # 3-year probability estimates with calibrate
  cal <- calibrate(rms_m, B=10, u=3)
  plot(cal, ylim = c(0,1))
  
  # K-fold validation with logistic model
  train_control <- trainControl(method = "cv", number = 10, savePredictions = TRUE)
  train_result <- train(as.formula(paste("DIABF ~ ", paste(c(m[["varselect"]], m[["varint"]]), collapse = "+"))), 
                        data = m[["data"]], trControl = train_control, method = "glm", family = "binomial")
  
  # Show the final model
  train_result$finalModel
  
  # Return result
  return(rms_m)
  
}
```

##### Full Model with 11 Variables 

This is the full model, 11 variables. This model includes BMI * treatment interaction (BMIGROUP*ASSIGN).

```{r model_11_var, echo=TRUE}
varselect <- c('AGEGROUP','RACE_ETH','ASSIGN','FamHx',
               'zWaistCircAve','zWTHRatio','BMIGROUP',
               'zG000','zHBA1', 'zlogTRIG','ztotalMETs')
varint <- c('ASSIGN*BMIGROUP')
labels <- c('40-44','45-49', '50-54', '55-59','60+',
            'African American', 'Hispanic, of any race', 'All other', 
            'Lifestyle', 'Metformin','FamHx', 
            'Waist circumference,Std','Waist:Hip Ratio,Std','BMI 30-35', 'BMI 35+', 'Fasting Glucose, Std', 'HbA1c, Std','Triglycerides (ln), Std','METS hr/week,Std',
            'Lifestyle * BMI 30-35', 'Metformin * BMI 30-35', 
            'Lifestyle * BMI 35+', 'Metformin * BMI 35+')
m_11_var <- run_model(varselect, varint, labels)
kable(m_11_var[["tidy_model"]], digits = 3)
kable(m_11_var[["tidy_model_exp"]], digits = 3)
m_11_var_rms <- validate_model(m_11_var)
```

\newpage

##### Proportional Hazards Assumption

This is to check proportional hazards assumptions.

```{r test_cph_assumption, echo = TRUE}
# Check each covariate (proportional hazards assumption)
res1 <- cox.zph(m_11_var[["model"]])
res1

# Plot residuals (Schoenfeld residuals vs. time)
plot(res1)
```

\newpage

##### Survival Plots

```{r survival_plots, echo = TRUE}

# Create new data frame and set DiabYes to 0 (no diabetes) and DIABT time to 3 (3 years)
newdf <- m_11_var[["data"]]
newdf$DIABT <- 3
newdf$DiabYes <- 0

# Calculate predicted probability of diabetes at 3 years based on cph model
m_11_var[["data"]]$survprob <- exp(-predict(m_11_var[["model"]], newdata = newdf, type = "expected"))

# Plot predicted probability by BMI graoup and treatment assignment
ggplot(m_11_var[["data"]], aes(y = survprob, x = BMIGROUP, fill = ASSIGN)) +
  geom_boxplot() +
  scale_y_continuous("3-year Predicted Probability of Survival\n(No Diabetes)") +
  scale_x_discrete() +
  labs(x= expression(paste("BMI group (kg/m"^{2},")"))) +
  scale_fill_manual(breaks = c("Placebo", "Lifestyle", "Metformin"), values = c("#BB1A58", "#CFCEE3", "#1C9DA9")) +
  theme_bw() +
  theme(text = element_text(size = 13), panel.grid.minor = element_blank(), panel.grid.major = element_blank(),
        legend.title=element_blank(), axis.text = element_text(size = 11))

# KM survival with treatment figure
surv <- m_11_var[["surv"]]
fitASSIGN <- survfit(surv ~ ASSIGN, data = m_11_var[["data"]])
summary(fitASSIGN)
plot(fitASSIGN, xlab = "Years", ylab = "Probability of Diabetes by Treatment")
ggsurvplot(fit = fitASSIGN, data = m_11_var[["data"]], xlab = "Years", ylab = "Probability of Diabetes Free", xlim = c(0, 3.25), 
           title = "Survival Curve by Treatment", 
           legend = "bottom", legend.title = "", legend.labs = c("Lifestyle", "Metformin", "Placebo"),
           palette = c("#BB1A58","#a0d42b", "#1C9DA9"))
```

\newpage

##### Observed Vs. Predicted Probability of Diabetes

Show plot comparing observed and predicted probabilities of diabetes.

```{r observed_vs_predicted, echo=TRUE}

# Convert predicted probabilities to factor of deciles
m_11_var[["data"]]$diabprobf <- factor(floor((1 - m_11_var[["data"]]$survprob) * 10),
                                levels = 0:9,
                                labels = c("0-9%","10-19%","20-29%","30-39%",
                                          "40-49%","50-59%","60-69%","70-79%",
                                          "80-89%","90-99%"))

# For each decile get std dev, std error, conf int
agg <- summarySE(m_11_var[["data"]], measurevar = "DiabYes", groupvars = "diabprobf")

# As NaNs produced (maybe only 1 case in a percentile) use only complete cases so some deciles may be missing
agg <- agg[complete.cases(agg), ]

# Create plot of observed vs. predicted probability of diabetes
ggplot(agg, aes(x = diabprobf, y = DiabYes)) +
       geom_point() +
       geom_errorbar(aes(ymin = pmax(DiabYes - ci, 0), ymax = pmin(DiabYes + ci, 1))) +
       scale_y_continuous("Observed prevalence of diabetes", breaks = seq(0, 1, .1)) +
       scale_x_discrete("Predicted 3-year probability of diabetes") +
       theme_bw()
```

##### Goodness of CPH Model Fit

```{r goodness_of_fit, echo=TRUE}

# Check proportional hazards assumption
ggcoxdiagnostics(m_11_var[["model"]], type="schoenfeld")

# Check nonlinearity
ggcoxdiagnostics(m_11_var[["model"]], type="martingale")

# Check influential observations
ggcoxdiagnostics(m_11_var[["model"]], type="deviance")
```

\newpage

##### Exploratory model with 6 clinical variables

6 variable model for exploration. Model was based on all parous women. Not included in manuscript.

```{r model_6_var, echo=TRUE}
varselect <- c('ASSIGN', 'zWTHRatio', 'BMIGROUP', 'zG000', 'zHBA1', 'zlogTRIG')
varint <- c('ASSIGN*BMIGROUP')
labels <- c('Lifestyle', 'Metformin',  'Waist:Hip Ratio, Std',
            'BMI 30-35', 'BMI 35+', 'Fasting Glucose, Std', 'HbA1c, Std', 'Triglycerides (ln), Std',
            'Lifestyle * BMI 30-35', 'Metformin * BMI 30-35', 
            'Lifestyle * BMI 35+', 'Metformin * BMI 35+')
m_6_var <- run_model(varselect, varint, labels)
kable(m_6_var[["tidy_model"]], digits = 3)
kable(m_6_var[["tidy_model_exp"]], digits = 3)
m_6_var_rms <- validate_model(m_6_var)
```

\newpage

##### Reduced model with 8 clinical variables

Model includes both WaistCircAve and WTHRatio with BMI * treatment interaction.

Model not in revised manuscript.

```{r reduced_8_var, echo=TRUE}
varselect <- c('RACE_ETH', 'ASSIGN','zWaistCircAve', 'zWTHRatio', 'BMIGROUP','zG000','zHBA1','zlogTRIG')
varint <- c('ASSIGN*BMIGROUP')
labels <- c('African American', 'Hispanic, of any race', 'All other',
            'Lifestyle', 'Metformin',
            'Waist Circumference Ratio,Std',
            'Waist Hip Ratio,Std',
            'BMI 30-35', 'BMI 35+', 
            'Fasting Glucose, Std', 'HbA1c, Std','Triglycerides (ln), Std',
            'Lifestyle * BMI 30-35', 'Metformin * BMI 30-35', 
            'Lifestyle * BMI 35+', 'Metformin * BMI 35+')
m_red_8_var <- run_model(varselect, varint, labels)
kable(m_red_8_var[["tidy_model"]], digits = 3)
kable(m_red_8_var[["tidy_model_exp"]], digits = 3)
m_red_8_var_rms <- validate_model(m_red_8_var)
```

\newpage

##### Reduced model with 7 clinical variables

Model includes Waist Circumference only (omit WTHRatio) + BMI * treatment interaction

Model not in revised manuscript.

```{r reduced_7_var, echo=TRUE}
varselect <- c('RACE_ETH', 'ASSIGN','zWaistCircAve', 'BMIGROUP','zG000','zHBA1','zlogTRIG')
varint <- c('ASSIGN*BMIGROUP')
labels <- c('African American', 'Hispanic, of any race', 'All other',
            'Lifestyle', 'Metformin',
            'Waist Circumference Ratio,Std',
            'BMI 30-35', 'BMI 35+', 
            'Fasting Glucose, Std', 'HbA1c, Std','Triglycerides (ln), Std',
            'Lifestyle * BMI 30-35', 'Metformin * BMI 30-35', 
            'Lifestyle * BMI 35+', 'Metformin * BMI 35+')
m_red_7_var <- run_model(varselect, varint, labels)
kable(m_red_7_var[["tidy_model"]], digits = 3)
kable(m_red_7_var[["tidy_model_exp"]], digits = 3)
m_red_7_var_rms <- validate_model(m_red_7_var)
```

\newpage

##### Reduced model with 6 clinical variables

Model without WaistCircAve and WTHRatio but includes BMI * treatment interaction

Includes race/ethnicity and different than prior "exploratory" 6-variable model.

```{r reduced_6_var, echo=TRUE}
varselect <- c('RACE_ETH', 'ASSIGN', 'BMIGROUP','zG000','zHBA1','zlogTRIG')
varint <- c('ASSIGN*BMIGROUP')
labels <- c('African American', 'Hispanic, of any race', 'All other',
            'Lifestyle', 'Metformin',
            'BMI 30-35', 'BMI 35+', 
            'Fasting Glucose, Std', 'HbA1c, Std','Triglycerides (ln), Std',
            'Lifestyle * BMI 30-35', 'Metformin * BMI 30-35', 
            'Lifestyle * BMI 35+', 'Metformin * BMI 35+')
m_red_6_var <- run_model(varselect, varint, labels)
kable(m_red_6_var[["tidy_model"]], digits = 3)
kable(m_red_6_var[["tidy_model_exp"]], digits = 3)
m_red_6_var_rms <- validate_model(m_red_6_var)
```

\newpage

##### Parsimonious model

Parsimonious model with only HbA1c, fasting glucose, but includes BMI * treatment interaction

```{r parsimonious_model, echo=TRUE}
varselect <- c('ASSIGN','BMIGROUP','zG000','zHBA1')
varint <- c('ASSIGN*BMIGROUP')
labels <- c('Lifestyle', 'Metformin',
            'BMI 30-35', 'BMI 35+', 
            'Fasting Glucose, Std','HbA1c, Std',
            'Lifestyle * BMI 30-35', 'Metformin *BMI 30-35', 
            'Lifestyle * BMI 35+', 'Metformin * BMI 35+')
m_par <- run_model(varselect, varint, labels)
kable(m_par[["tidy_model"]], digits = 3)
kable(m_par[["tidy_model_exp"]], digits = 3)
m_par_rms <- validate_model(m_par)
```

\newpage

##### Survival Plots

```{r survival_plots, echo = TRUE}

# Create new data frame and set DiabYes to 0 (no diabetes) and DIABT time to 3 (3 years)
newdf <- m_par[["data"]]
newdf$DIABT <- 3
newdf$DiabYes <- 0

# Calculate predicted probability of diabetes at 3 years based on cph model
m_par[["data"]]$survprob <- exp(-predict(m_par[["model"]], newdata = newdf, type = "expected"))

# KM survival with treatment figure
surv <- m_11_var[["surv"]]
fitASSIGN <- survfit(surv ~ ASSIGN, data = m_11_var[["data"]])
summary(fitASSIGN)
plot(fitASSIGN, xlab = "Years", ylab = "Probability of Diabetes by Treatment")
ggsurvplot(fit = fitASSIGN, data = m_11_var[["data"]], xlab = "Years", ylab = "Probability of Diabetes Free", xlim = c(0, 3.25), 
           title = "Survival Curve by Treatment", 
           legend = "bottom", legend.title = "", legend.labs = c("Lifestyle", "Metformin", "Placebo"),
           palette = c("#BB1A58","#a0d42b", "#1C9DA9"))
```
\newpage

##### Comparison of model performance

The lrtest function does likelihood ratio tests for two nested models, from fits that have stats components with "Model L.R." values. For models such as psm, survreg, ols, lm which have scale parameters, it is assumed that scale parameter for the smaller model is fixed at the estimate from the larger model (see the example in rms package).

All combinations of the above models are compared.

```{r echo=TRUE}
# Comparisons of nested models
model_lst <- list(m_6_var_rms, m_11_var_rms, m_red_8_var_rms, m_red_7_var_rms, m_red_6_var_rms, m_par_rms)
lapply(model_lst,
  function(x) {
    lapply(model_lst, function(y) {
      if (!identical(x, y))
        rms::lrtest(x, y)
    })
  }
)
```

\newpage

##### ROC Curves

Survival ROC curve using example from:

https://datascienceplus.com/time-dependent-roc-for-survival-prediction-models-in-r/

For supplement in manuscript.

```{r roc_curves, echo = FALSE}

# Calculate predicted probability of diabetes at 3 years based on cph model
m_11_var[["data"]]$survprob <- predict(m_11_var[["model"]], type = "lp")
m_par[["data"]]$survprob <- predict(m_par[["model"]], type = "lp")

fullROC <-  survivalROC(Stime = m_11_var[["data"]]$DIABT, 
            status = ifelse(m_11_var[["data"]]$DIABF == 'Yes', 1, 0),
            marker = m_11_var[["data"]]$survprob, 
            method = "KM", 
            predict.time = 3)
fullROC.df <- data.frame(FP = fullROC$FP, TP = fullROC$TP, cv = fullROC$cut.values, 
                         model = paste0("Full Model (AUC ", round(fullROC$AUC, 3), ")"))

parROC <-   survivalROC(Stime = m_par[["data"]]$DIABT, 
            status = ifelse(m_par[["data"]]$DIABF == 'Yes', 1, 0),
            marker = m_par[["data"]]$survprob, 
            method = "KM", 
            predict.time = 3)
parROC.df <- data.frame(FP = parROC$FP, TP = parROC$TP, cv = parROC$cut.values, 
                        model = paste0("Parsimonious Model (AUC ", round(parROC$AUC, 3), ")"))

all_ROC.df <- bind_rows(fullROC.df, parROC.df)

# Draw the plot of TP vs. FP
ggplot (data = all_ROC.df, mapping = aes(x = FP, y = TP, label = cv, color = model)) +
  geom_roc(labels = FALSE, stat = "identity") +
  style_roc(theme = theme_gray) +
  labs(title = "Survival ROC - Full and Parsimonious Models") +
  theme(legend.title = element_blank())
```

##### END